The Impact of Home Advantage, Goals Scored, Shots on Target on Primier League Outcomes

### 2024-04-17 paper/paper.qmd
updates on discussion section in paper, and reference

>>>>

# Discussion {#sec-discussion}

## Home Advantage

Home advantage has been a prominent feature of team sports and football is no exception to this. Regardless of division level, the home advantage is worth over 0.5 of a goal on average [@Clarke]. It is evident that home advantage, which plays a crucial role in football, especially in tournaments, indeed exists. However, the extent of its impacts varies across teams. Our analysis reveals that while some teams play better at home, others do not capitalize on it as expected. Although the regression coefficients did not fully capture the home advantage, the disparities in the winning rates between home teams and away teams in @tbl-home and @tbl-away suggest that in the Premier League, home advantage exists and is quite important. Factors such as home crowd's support on player's performance and referee's decision-making, adverse effects of travelling, jet lag and fatigue, and psychologically defensive reaction to protect territory have been identified to contribute to home advantage [@Goumas]. Past research suggests that referees can be influenced by social pressure from the home supporters [@Endrich]. This suggests that many factors beyond the physical venue can affect home advantage. One of the implications of the findings is that teams should do their best to win in their home match. Considering its benefits, a draw or a defeat at home is crucial especially for those competing for titles or avoiding relegation. This further suggests that the teams in the Champions League or Europa League should design their tactics to aggressively secure home wins to maximize this advantage.

## Goals scored and The Winning Rate

The relationship between the winning rate and goals scored is one of the most straightforward metrics in football. Past studies suggest that winning teams show stronger results in goals scored than losing and drawing teams [@Penas]. Our findings coincide with this view, which is shown effectively in @fig-goal. It clearly shows that an increase in the average number of goals scored by a team is significantly correlated with an increase in their winning rate. This correlation goes back to the basic strategy of football: to win games by outscoring the opponent. The data shows that teams with higher average goals scored are not just winning more frequently but are often dominating the league. These teams are Man City, Liverpool and Arsenal from the 2023-24 season match data. This reinforces the need for teams to develop strong offensive capabilities as a critical component for their match strategy. This further leads to team tactics and training. To play better football offensively, you need to have higher shot quality, better positioning, better ball possession and build-up play into the opposition box to create goal scoring opportunities. Some very effective tactical managers like Pep Guardiola of Man City, Mikel Arteta of Arsenal, Jürgen Klopp of Liverpool, Xabi Alonso of Leverkusen and Roberto De Zerbi of Brighton & Hove Albion are dominating the world football in terms of new tactics. They show progressive tactics in their games, and other teams try to find ways to copy or defend their tactics. In the future, tactical planning and the execution will be more important in modern football where players are now playing like a unit of a perfect mechanism compared to the past where an individual player's talent and skills were more important. Strikers and midfielders with creativity and work rate to consistently contribute to create more scoring opportunities will be more valuable to their teams. Defenders with great ball skills and passing abilities will be more valuable in the future.

## Shots on Target and The Winning Rate

Among scholars and professionals, shooting in football has been a subject of keen interest and analysis [@Rathke]. Previous research suggest thats winning teams made more shots and shots on target than the opponents [@Penas]. @fig-shot confirms a positive correlation between the number of shots on target and the winning rate, suggesting that increasing shots on target can increase a team's probability of winning. This is in line with the relationship between goals scored and the winning rate. The better ability to break down the opposition's defense and create scoring opportunities, the better chance of winning in football. This underscores the importance of the strategy of attacking. It is not just about quantity but the quality of shots because to shoot on target, it requires both precision and power in their attacks. Detailed training should focus on shooting accuracy, decision-making in the final third, and tactical movements that can lead to the creation of shooting opportunities. This metric can also be a evaluation tool for individual players especially for forwards, wingers and attacking midfielders. Players who have higher shots on target ratio are the ones more likely to convert chances into goals.

## Weaknesses and Implications for Future Research

Despite these findings, the model, while insightful, relies on oversimplification to represent the complexity of football. Factors such as players' skills, team cohesion, fitness levels, mentality, tactical strategies: formation, in-game adjustments, pitch conditions, injuries, referee decisions are not captured but could significantly affect the outcomes of matches.

The negative coefficient associated with the proportion of home games suggests that this coefficient does not accurately capture home advantage given the 85% of teams performing better at home in our data. In the future, the study of home advantage could look into more specific aspects like crowd, referee's decision-making in home versus away games or psychological factors.

The study lacks in differentiation between different types of goals such as open play, set-pieces, penalties and solo, each of which might have different implications on the effect of goals scored and the winning rates. Looking into these details may provide better understanding. Integration of the impact of defensive tactics and pressure, assists, player movement, different types of passes can help understand the build-up play that leads to more goals. This can provide a more comprehensive view of how teams can increase their scoring opportunities to win the match.

\newpage

# References
